---
title: "Ruth Gates Analysis and Figures Part 1: Stats"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load all of the needed libraries.

```{r}
library(devtools)

#install.packages("glmmTMB",type='source') #will need to install development tools and gfortran
#install.packages("wesanderson")
#install.packages("lme4", type = "source")
#install.packages("vcfR")

# Graphical tools :
library(ggplot2)
library(cowplot)
library(gridExtra)
#library(nationalparkcolors)

# Data Wrangling :
library(dplyr)
library(tidyr)
library(forcats)
library(car)
library(vcfR)

# Stats Packages :
library(lme4)
library(lsmeans)
library(lmerTest)
library(MuMIn)
library(Matrix)
library(glmmTMB)
library(bbmle)
library(DHARMa) 
## Learning DHARMa - https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
library(multcomp)
library(betareg)
library(emmeans)
library(multcompView)

#pretty stats table function
library(xtable)
pxt <- function(x,title) {
cat(sprintf("{\n\n\\textbf{%s}\n\\ \\\\\\vspace{2pt}\\ \\\\\n",title))
13
print(xtable(x), floating=FALSE); cat("\n\n")
cat("\\ \\\\\\vspace{5pt}\\ \\\\\n")
}

# Survival analysis :
library(survival) 
library(coxme)
library(survminer) #for plots
library(wesanderson)

```

Load in the quality filtered trait data and integrated symbiont type dataset

## Import the final QC-ed datafile for subsequent statistical analysis

```{r}

traitsFiltered_SymTypeSubset<-readRDS(file="~/Dropbox/CarlsLab/ResearchProjects/NOAA_RuthGates/TraitData/traitsFiltered_SymTypeSubset_PostHocFam.rds")

Nori<-traitsFiltered_SymTypeSubset %>%
    group_by(Genet) %>%
    count(Origin_PostHoc) 

summary(Nori)

```

## Now for some Trait statistics

Note, Racks are nested within tanks as random effects and genets are within families. To focus on cross effects Must create data subset with 3 high replication Batch Cross Origins to ask about diffs in variance among crosses

```{r}

traitsAllCrossSym <- traitsFiltered_SymTypeSubset %>%
  filter(Origin_PostHoc=="E-B Batch Cross A" | Origin_PostHoc=="E-B Batch Cross B" | Origin_PostHoc=="E-B Batch Cross C" | Origin_PostHoc=="LK3 x LK10 Biparental Cross" | Origin_PostHoc=="E-H-SI Batch Cross") %>%
  droplevels()

summary(traitsAllCrossSym)

#create dummy variable for parsing of any interaction effects
traitsAllCrossSym$OriTrt<-with(traitsAllCrossSym,factor(Origin_PostHoc:Treatment))


```

Exploratory plots suggest difference in symbiont type among families - will need to test for and account for this in subsequent trait models. Symbiont data are almost binary with many near 0 and near 1...need beta regression

```{r}

hist(traitsAllCrossSym$X.D,breaks=30) #not horrible actually but behave almost like a binomial
hist(traitsAllCrossSym$X.A,breaks=30) #Is prop D is more informative bc more ones than zeros??
hist(traitsAllCrossSym$X.C,breaks=30) #C is really non-informative and quite infrequent

#create binary to distinguish zero from non-zero max growth for exploratory plots...prop A would be largely inverse but more zeros

traitsAllCrossSym$D_nonzero<-ifelse((traitsAllCrossSym$X.D > 0 & !is.na(traitsAllCrossSym$X.D)), 1, 0)

ggplot(traitsAllCrossSym, aes(Treatment, X.D, colour = as.factor(D_nonzero))) + geom_jitter() #small treatment effect

ggplot(traitsAllCrossSym, aes(Origin_PostHoc, X.D, colour = as.factor(D_nonzero))) + geom_jitter() #very much origin effect

ggplot(traitsAllCrossSym, aes(Treatment, X.D, colour = as.factor(D_nonzero))) + geom_jitter()+facet_wrap(~Origin_PostHoc) #not really interaction

ggplot(traitsAllCrossSym, aes(Genet, X.D, colour = as.factor(D_nonzero))) + geom_jitter() #genet effect somewhat noisy...origin seems bigger driver

ggplot(traitsAllCrossSym, aes(Raceway, X.D, colour = as.factor(D_nonzero))) + geom_jitter() #raceway effect not really noticeable

#lets try fitting with plain beta regression - for values bounded between 0 and 1
BR_d<-betareg(X.D ~ Origin_PostHoc+Treatment+Origin_PostHoc:Treatment,data=traitsAllCrossSym)
summary(BR_d) #only fixed effects
AIC(BR_d)
summary(BR_d)$pseudo.r.squared #really not a lot of variance explained
joint_tests(BR_d) #mega effect of origin, minor treatment, none for interaction really
plot(fitted(BR_d),
     residuals(BR_d)) #ugly...why so many bins?
plot(BR_d)

marginal = emmeans(BR_d,
                   ~ Origin_PostHoc) 

pairs(marginal,
      adjust="tukey")


Sum = cld(marginal,
          alpha   = 0.05,
          Letters = letters,         ###  Use lowercase letters for .group
          adjust  = "tukey")         ###  Tukey-adjusted comparisons


Sum


traitsAllCrossSym$GenetOri<-with(traitsAllCrossSym,factor(Origin_PostHoc:Genet))
traitsAllCrossSym$GenetOriR = relevel(traitsAllCrossSym$GenetOri, ref = "E-H-SI Batch Cross:2") #D dominated ref genet

#curious - what if we drop interaction to simplify model and add genet effect- better fit?
BR_d2<-betareg(X.D ~ GenetOriR+Treatment,data=traitsAllCrossSym)
summary(BR_d2)
AIC(BR_d2)
summary(BR_d2)$pseudo.r.squared #MUCH better fit

plot(fitted(BR_d2),
     residuals(BR_d2)) #better
plot(BR_d2) #best model

marginal2 = emmeans(BR_d2,
                   ~ GenetOriR)

tabD2<-emmeans(BR_d2,~ Treatment)
tabD2
summary(tabD2)[2,2]-summary(tabD2)[1,2] #take difference in means, heat - ctrl

AIC(BR_d,BR_d2)

#Number of EB-A genets = 45
length(grep("GenetOriRE-B Batch Cross A",rownames(summary(BR_d2)$coefficients$mean)))

#how many significant - EB-A? 12
length(grep("GenetOriRE-B Batch Cross A",names(which(summary(BR_d2)$coefficients$mean[,4]<0.05)))) 

#Number of EB-B genets = 17
length(grep("GenetOriRE-B Batch Cross B",rownames(summary(BR_d2)$coefficients$mean)))

#how many significant - EB-B? 7
length(grep("GenetOriRE-B Batch Cross B",names(which(summary(BR_d2)$coefficients$mean[,4]<0.05)))) 

#Number of EB-C genets = 18
length(grep("GenetOriRE-B Batch Cross C",rownames(summary(BR_d2)$coefficients$mean)))

#how many significant - EB-C? 6
length(grep("GenetOriRE-B Batch Cross C",names(which(summary(BR_d2)$coefficients$mean[,4]<0.05)))) 

#Number of LK genets = 42
length(grep("GenetOriRLK3",rownames(summary(BR_d2)$coefficients$mean)))

#how many significant? 1
length(grep("GenetOriRLK3",names(which(summary(BR_d2)$coefficients$mean[,4]<0.05)))) 


#Number of EHSI genets = 23+1 for ref = 24
length(grep("GenetOriRE-H",rownames(summary(BR_d2)$coefficients$mean)))

#how many significant? 12
length(grep("GenetOriRE-H",names(which(summary(BR_d2)$coefficients$mean[,4]<0.05)))) 


#NOW does proportion Symbiodinium mirror Durusdinium results? it should
BR_a<-betareg(X.A ~ Origin_PostHoc+Treatment+Origin_PostHoc:Treatment,data=traitsAllCrossSym)
summary(BR_a)
summary(BR_a)$pseudo.r.squared #same same
joint_tests(BR_a) #mega effect of origin, minor treatment, none for interaction 

plot(fitted(BR_a),
     residuals(BR_a)) #ugly...why so many bins?
plot(BR_a)

marginal = emmeans(BR_a,
                   ~ Origin_PostHoc)

pairs(marginal,
      adjust="tukey")


Sum = cld(marginal,
          alpha   = 0.05,
          Letters = letters,         ###  Use lowercase letters for .group
          adjust  = "tukey")         ###  Tukey-adjusted comparisons


Sum

## now with genet nested within origin
BR_a2<-betareg(X.A ~ GenetOriR+Treatment,data=traitsAllCrossSym)
summary(BR_a2)
summary(BR_a2)$pseudo.r.squared #same same, much better

plot(fitted(BR_a2),
     residuals(BR_a2)) #better
plot(BR_a2)

tabA2<-emmeans(BR_a2,~ Treatment)
tabA2
summary(tabA2)[2,2]-summary(tabA2)[1,2] #take difference in means, heat - ctrl

AIC(BR_a,BR_a2) #again, includign genet makes better model



#main conclusion here is that the Proportion of Durusdinium varies as a function of family and treatment, but not the interaction. Durusdinium based trade-offs can influence growth/survival - need to account for this when modeling effects of cross type and treatment on sym photophys, growth and survival

```

S:H Cell Ratio

```{r}

hist(traitsAllCrossSym$SH_T, breaks=30)
summary(traitsAllCrossSym$SH_T) #more strongly skewed data...

ggplot(traitsAllCrossSym, aes(Treatment, SH_T)) + geom_jitter() #mostly just some strong outliers

ggplot(traitsAllCrossSym, aes(Origin_PostHoc, SH_T)) + geom_jitter() #same

ggplot(traitsAllCrossSym, aes(Treatment, SH_T)) + geom_jitter()+facet_wrap(~Origin_PostHoc) #not really interaction

ggplot(traitsAllCrossSym, aes(Genet, SH_T)) + geom_jitter() #genet effect somewhat noisy...origin seems bigger driver

ggplot(traitsAllCrossSym, aes(Raceway, SH_T)) + geom_jitter() #raceway effect not really noticeable

#what if we just log transform and run lmer
hist(log10(traitsAllCrossSym$SH_T), breaks=30) #now looks nice...suspect will work

#traitsCross_SAclean<-traitsCross[complete.cases(traitsCross$SA_dif_cor_log),]

SH_T_model <- lmer(log10(traitsAllCrossSym$SH_T) ~ X.D+Origin_PostHoc*Treatment  + (1|Genet) + (1|Raceway), data=traitsAllCrossSym)

qqPlot(residuals(SH_T_model),xlab="Theoretical Quantiles",ylab="Observed Quantiles") #looks very nice
plot(SH_T_model,which=1) #feeling pretty good about this

shRes <- simulateResiduals(fittedModel = SH_T_model)
plot(shRes) #not terrible

summary(SH_T_model) 

anova(SH_T_model,ddf="Kenward-Roger") 

traitsAllCrossSym %>% 
  mutate(log_SHT=log10(SH_T)) %>%
  ggplot(aes(x = Origin_PostHoc, y = log_SHT)) + 
  geom_boxplot() #verify origin effect

rand(SH_T_model) #genet non-sig, raceway very sig (possibly conflated with treatment)

Ori <- glht(SH_T_model, linfct = mcp(Origin_PostHoc = "Tukey"))
summary(Ori) 

```

Photophysiology

```{r}

############### Fv/Fm

hist(traitsAllCrossSym$FvFm_PropDif, breaks=20) #looks great

FvFm_model <- lmer(FvFm_PropDif ~ X.D+Origin_PostHoc*Treatment  + (1|Genet) + (1|Raceway), data=traitsAllCrossSym) #default is gaussian(identity)

resFvFm <- simulateResiduals(fittedModel = FvFm_model)
plot(resFvFm) #pretty okay


summary(FvFm_model) #what if we decide this model is good enough...

anova(FvFm_model,ddf="Kenward-Roger") #lots going on, effect of D and interaction

emmeans(FvFm_model,~ Origin_PostHoc)


rand(FvFm_model) #strong random effects as well, raceway bigger than genet

#need to rerun with dummy interaction variable to get proper contrasts

FvFm_model2 <- lmer(FvFm_PropDif ~ X.D+OriTrt  + (1|Genet) + (1|Raceway), data=traitsAllCrossSym) #default is gaussian(identity)

emmeans(FvFm_model2,~ OriTrt)

resFvFm2 <- simulateResiduals(fittedModel = FvFm_model2)
plot(resFvFm2) #same

OT <- glht(FvFm_model2, linfct = mcp(OriTrt = "Tukey"))
summary(OT) 

#EB batch A shows treatment effect
#also baseline diff in control conditions
#rest of pairwise are silly

```

## Survival Data

```{r}

names(traitsAllCrossSym)

traitsAllCrossSym %>%
    group_by(Origin_PostHoc, Treatment) %>%
    summarise (propDead = mean(Mortality),meanTOD = mean(TOD)) #No mortality in control; only mortality in heat... variation among crosses?

traitsAllCrossSym$Genet<-relevel(traitsAllCrossSym$Genet,ref="4") #relevel with best surviving genet as reference
traitsAllCrossSym$Origin_PostHoc<-relevel(traitsAllCrossSym$Origin_PostHoc,ref="E-B Batch Cross B") #relevel with best surviving cross as reference

summary(traitsAllCrossSym$TOD)

heats<-traitsAllCrossSym %>%
  filter(Treatment=="Heat") %>%
  droplevels()

ctrls<-traitsAllCrossSym %>%
  filter(Treatment=="Control") %>%
  droplevels()

summary(heats$TOD)
summary(ctrls$TOD) #coding error - different max survival time artifact - need to make identical - solved

#survival by cross origin and treatment
iter<-coxme.control(iter.max=500)

died=Surv(traitsAllCrossSym$TOD, traitsAllCrossSym$Mortality)

table(died) 

survObT<-coxme(died~Origin_PostHoc*Treatment+(1|Raceway), data=traitsAllCrossSym,control=iter) #does not converge bc no variance for ctrls

summary(survObT) #yes, cannot estimate p-value; therefore will run as simpler model first

survT<-coxme(died~Treatment+(1|Raceway), data=traitsAllCrossSym,control=iter) #does not converge bc no variance for ctrls

summary(survT)#yeah, blows up bc treatment coefficient is insane - will run subset - just heat-treated

diedH=Surv(heats$TOD, heats$Mortality) 

table(diedH) 

survO<-coxme(diedH~Origin_PostHoc+(1|Raceway), data=heats,control=iter) #def origin effect

survO

summary(survO) #sig higher mortality risk in E-B A and E-H-SI batch, but need to control for sym type

###what if we now add in an effect of proportion Durusdinium? Might not be possible if deads don't have Durusdinium

heats %>%
  filter(Mortality>0.99)#there is data on sym type here...

survOD<-coxme(diedH~Origin_PostHoc+X.D+(1|Raceway), data=heats,control=iter) #def origin effect

survOD

summary(survOD) #OOOOHHHHH Durusdinium reduces mortality risk, but still have Origin effect with E-H-SI and EB-A being higher mortality risk than E-B B/C and LK biparental

heats$Origin_PostHoc<-relevel(heats$Origin_PostHoc,ref="LK3 x LK10 Biparental Cross") #relevel to get pairwise origin difs

survOD2<-coxme(diedH~Origin_PostHoc+X.D+(1|Raceway), data=heats,control=iter) #def origin effect

summary(survOD2) #E-H-SI and EB A also sig worse than LK and E-B B and E-B C 

heats$Origin_PostHoc<-relevel(heats$Origin_PostHoc,ref="E-B Batch Cross C") #relevel to get pairwise origin difs

survOD3<-coxme(diedH~Origin_PostHoc+X.D+(1|Raceway), data=heats,control=iter) #def origin effect

summary(survOD3) #E-H-SI and EB A also sig worse than LK and E-B B and E-B C 


#BUT need to think about potential interaction effects since Prop D differs by Origin - lets make artificial bins to try to calculate
head(heats)

hist(heats$X.D,breaks=30) #lots of D and A dominated, and transitional

heats$D_bins<-cut(heats$X.D,
                  breaks=c(0,0.1,0.9,1),
                  labels=c('negligible','transitional','dominant')) #defining transitional as more than 10% and less than 90%

heats$D_bins<-relevel(heats$D_bins,ref="dominant") #relevel to get pairwise

heats$Origin_PostHoc<-relevel(heats$Origin_PostHoc,ref="E-B Batch Cross A") #relevel to get pairwise origin difs


survOCatD<-coxme(diedH~Origin_PostHoc*D_bins+(1|Raceway), data=heats,control=iter) #interaction model not successful

survOCatD<-coxme(diedH~Origin_PostHoc+D_bins+(1|Raceway), data=heats,control=iter) #will do additive

survOCatD

summary(survOCatD) #E-H-SI and almost EB-A sig higher mortality than EB-B, sym A dominant is def worse

heats$D_bins<-relevel(heats$D_bins,ref="transitional") #relevel to get pairwise

survOCatD2<-coxme(diedH~Origin_PostHoc+D_bins+(1|Raceway), data=heats,control=iter) #will do additive

summary(survOCatD2) #E-H-SI and almost EB-A sig higher mortality than EB-B, and any amount of D helps; less than 10% is higher mort risk than transitional 

################### Kaplan-Meier survival curves

###################### try to visualize additive effects of %D and family Origin 

heats$D_bins<-relevel(heats$D_bins,ref="dominant") #relevel to get pairwise
heats$Origin_PostHoc<-relevel(heats$Origin_PostHoc,ref="LK3 x LK10 Biparental Cross") #relevel to get plot colors correct

heats$OriDbins<-with(heats,factor(Origin_PostHoc:D_bins))

levels(heats$OriDbins)

surPlotOD = survfit(diedH ~ OriDbins, data=heats)

summary(heats)

#what if we facet by D.bins - YES much better

dev.new()
ggsurvplot_facet(
  surPlotOD,
  data = heats,
  facet.by = "D_bins",
  size = 1,                 # change line size
  palette =
    c("#999999","tan","tan","tan","goldenrod1","goldenrod1","goldenrod1","yellow2","yellow2","yellow2","#CC6677","#CC6677","#CC6677"),# custom color palettes
  conf.int = FALSE,          # Add confidence interval
#  pval = TRUE,              # Add p-value
#  risk.table = TRUE,        # Add risk table
#  risk.table.col = "strata",# Risk table color by groups
#  legend.labs =
#    c("None", "Transitional","Dominant"),    # Change legend labels
#  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw()      # Change ggplot2 theme
)
#dev.off()

#"#DDCC77","lightgoldenrod3","burlywood2"

###finally, what if we export genet specific hazard ratio in heat, then in GWAS control for relatedness and Durusdinium dominance

names(traitsFiltered_SymTypeSubset)

traitsFiltered_SymTypeSubset %>%
    group_by(Origin_PostHoc, Treatment) %>%
    summarise (propDead = mean(Mortality),meanTOD = mean(TOD)) #No mortality in control; only mortality in heat... 

traitsFiltered_SymTypeSubset$Genet<-relevel(traitsFiltered_SymTypeSubset$Genet,ref="4") #relevel with best surviving genet as reference

summary(traitsFiltered_SymTypeSubset$TOD)

heats<-traitsFiltered_SymTypeSubset %>%
  filter(Treatment=="Heat") %>%
  droplevels()

#survival by cross origin and treatment
iter<-coxme.control(iter.max=500)

diedH=Surv(heats$TOD, heats$Mortality) 

table(diedH) 

survG<-coxme(diedH~Genet+(1|Raceway), data=heats,control=iter) #coefficients might be crazy

summary(survG)

survG_fixef<-fixef(survG)

survG_HR <- exp(survG_fixef)

survG_HR <- as.data.frame(survG_HR)

summary(survG_HR)

head(survG_HR)

hist(survG_HR$survG_HR)

sort(survG_HR$survG_HR)

survG_HR$HR_bin<-survG_HR$survG_HR

survG_HR$HR_bin<-cut(survG_HR$survG_HR,
                  breaks=c(0,1,2.505775e+15),
                  labels=c(1,0))

survG_HR

write.csv(survG_HR,file="GenetSurvivalinHeat_coxmeHazardRatioBins.csv",quote=FALSE,row.names=TRUE)

#NOTE its equivalent to binning by no mortality vs any mortality

```

Now, test for differences in surface area growth by cross origin and treatment. Data are zero-inflated - solution: use zero inflated model, but first lets explore distribution of response by predictors

```{r}

#create binary to distinguish zero from non-zero max growth for exploratory plots

traitsAllCrossSym$SA_nonzero<-ifelse(traitsAllCrossSym$SA_dif_cor > 0, 1, 0)

ggplot(traitsAllCrossSym, aes(Treatment, SA_dif_cor, colour = as.factor(SA_nonzero))) + geom_jitter() #does appear to be treatment effect, more zeros in heat

ggplot(traitsAllCrossSym, aes(Origin_PostHoc, SA_dif_cor, colour = as.factor(SA_nonzero))) + geom_jitter() #does appear to be origin effect, more in EB-A?

ggplot(traitsAllCrossSym, aes(Treatment, SA_dif_cor, colour = as.factor(SA_nonzero))) + geom_jitter()+facet_wrap(~Origin_PostHoc) #not really interaction?

traitsAllCrossSym %>%
  filter(SA_nonzero==1) %>%
  ggplot(aes(Treatment, SA_dif_cor, colour = as.factor(Treatment))) + geom_boxplot()+facet_wrap(~Origin_PostHoc)

ggplot(traitsAllCrossSym, aes(Genet, SA_dif_cor, colour = as.factor(SA_nonzero))) + geom_jitter() #genet effect somewhat noisy...origin seems bigger driver

ggplot(traitsAllCrossSym, aes(Raceway, SA_dif_cor, colour = as.factor(SA_nonzero))) + geom_jitter() #raceway effect present, but stratified by treatment

### really seems like treatment is the thing driving low SA growth either 0 or continuous. Would prefer zero-inflated model...

# try glmmTMB with zero inflated gamma distribution
#Note- cannot include additional random effect of Genet, get infinite residuals in simulations - possibly bc genet is nested within origin??
SA_dif_model0 <- glmmTMB(SA_dif_cor ~ X.D+Origin_PostHoc+Treatment + (1|Raceway), ziformula = ~Origin_PostHoc+Treatment , family = "ziGamma", data = traitsAllCrossSym)

res0 <- simulateResiduals(fittedModel = SA_dif_model0) #check fits with DHARMa
plot(res0) #everything is looking pretty darn good now
testZeroInflation(res0) #all good!

# add in interaction
SA_dif_model1 <- glmmTMB(SA_dif_cor ~ X.D+Origin_PostHoc*Treatment + (1|Raceway), ziformula = ~Treatment+Origin_PostHoc , family = "ziGamma", data = traitsAllCrossSym)

res1 <- simulateResiduals(fittedModel = SA_dif_model1)
plot(res1) #everything is looking pretty darn good now
testZeroInflation(res1) #solved! totally good

# what about interaction in zero inflated mode?
SA_dif_model2 <- glmmTMB(SA_dif_cor ~ X.D+Origin_PostHoc*Treatment + (1|Raceway), ziformula = ~Treatment*Origin_PostHoc , family = "ziGamma", data = traitsAllCrossSym) #model is too complicated - singular convergence; missing levels for zero inflated interaction

# alternatively accounting for symbiont type with additive zi interaction?
SA_dif_model3 <- glmmTMB(SA_dif_cor ~ X.D+Origin_PostHoc*Treatment + (1|Raceway), ziformula = ~X.D+Treatment+Origin_PostHoc , family = "ziGamma", data = traitsAllCrossSym) #converges

res3 <- simulateResiduals(fittedModel = SA_dif_model3)
plot(res3) #everything is looking pretty darn good now
testZeroInflation(res3) #solved! totally good


###okay which model 
AICtab(SA_dif_model0,SA_dif_model1,SA_dif_model2,SA_dif_model3) #more complex models better but fail to converge, lets go with model3

SA_dif_model3 
summary(SA_dif_model3) #more complex models fail to converge
Anova(SA_dif_model3) #NOTE this is Wald chi2 test (NO LRT nor F test), only apply to fixed effects of conditional model

#the conditional output represents the zero portion (or a logistic regression) - the zero inflated output represents a "mixture" model of the two distributions - one for the zero subgroup and one for the non-zero subgroup
#The ziformula is a gamma model turned up side down, it tells you which of your predictors increase the probability of no growth vs some growth. The conditional model tells you which predictors influence the max growth, for those cases where some growth is evident. 


#need to create dummy variable to get proper pairwise comparisons for interaction, then rerun stats and do Tukeys


SA_dif_model3.1 <- glmmTMB(SA_dif_cor ~ X.D+OriTrt + (1|Raceway), ziformula = ~X.D+Treatment+Origin_PostHoc , family = "ziGamma", data = traitsAllCrossSym)

res3.1 <- simulateResiduals(fittedModel = SA_dif_model3.1)
plot(res3.1) #same same
testZeroInflation(res3.1) #good

saI <- glht(SA_dif_model3.1, linfct = mcp(OriTrt = "Tukey"))
summary(saI) #note that not all contrasts are meaningful, so this is a bit overly conservative, but 

#now lets print table of coefficients
sum<-summary(SA_dif_model3)

pxt(lme4::formatVC(sum$varcor$cond),"Random effect variances")
pxt(coef(sum)$cond,"Conditional fixed effects")
pxt(coef(sum)$zi,"Conditional zero-inflation effects")

#upload code to: https://latexbase.com/d/9eb8490b-dd46-48b1-89ea-258aebec6647 to get tables


```

Partial Mortality is also zero inflated - same model structure as above

```{r}

hist(traitsAllCrossSym$PartMort, breaks=20)

#create binary to distinguish zero from non-zero mortality

traitsAllCrossSym$PM_nonzero<-ifelse(traitsAllCrossSym$PartMort > 0, 1, 0)

ggplot(traitsAllCrossSym, aes(Treatment, PartMort, colour = as.factor(PM_nonzero))) + geom_jitter() #def treatment effect

ggplot(traitsAllCrossSym, aes(Origin_PostHoc, PartMort, colour = as.factor(PM_nonzero))) + geom_jitter() #origin effect not clear

ggplot(traitsAllCrossSym, aes(Treatment, PartMort, colour = as.factor(PM_nonzero))) + geom_jitter()+facet_wrap(~Origin_PostHoc) #possibly interaction

ggplot(traitsAllCrossSym, aes(Genet, PartMort, colour = as.factor(PM_nonzero))) + geom_jitter() #genet effect again noisy

ggplot(traitsAllCrossSym, aes(Raceway, PartMort, colour = as.factor(PM_nonzero))) + geom_jitter() #raceway effect present, but stratified by treatment

# again glmmTMB with zero inflated gamma distribution - include Prop D as covariate just to control
#Might want to consider interaction in zero inflation model...
PM_dif_model0 <- glmmTMB(PartMort ~ X.D+Origin_PostHoc+Treatment + (1|Raceway), ziformula = ~X.D+Origin_PostHoc+Treatment , family = "ziGamma", data = traitsAllCrossSym)

res0pm <- simulateResiduals(fittedModel = PM_dif_model0)
plot(res0pm) #niiiiice
testZeroInflation(res0pm)

PM_dif_model0
summary(PM_dif_model0)
Anova(PM_dif_model0) #NOTE this is Wald chi2 test (NO LRT nor F test), only applies to fixed effects of conditional model

#probability of no partial mortality strongly influenced by treatment - 4x more likely to observe no partial mortality in control than in heat
#WHEN partial mortality is evident, there is a strong effect of Durusdinium, but also origin and treatment

# testing interaction model
PM_dif_model1 <- glmmTMB(PartMort ~ X.D+Origin_PostHoc*Treatment + (1|Raceway), ziformula = ~X.D+Treatment*Origin_PostHoc , family = "ziGamma", data = traitsAllCrossSym) 

#can't handle interaction...issue is likely complete separation? no instances of partial mortality in Biparentals in control.
ggplot(traitsAllCrossSym, aes(Treatment, PartMort, colour = as.factor(PM_nonzero))) + geom_jitter()+facet_wrap(~Origin_PostHoc)

#in this case, run reduced model of heat-treated ramets only to try to investigate potential origin effects
heats<-traitsAllCrossSym %>%
  filter(Treatment=="Heat") %>%
  droplevels()

PM_heats_model0 <- glmmTMB(PartMort ~ X.D+Origin_PostHoc + (1|Raceway), ziformula = ~X.D+Origin_PostHoc , family = "ziGamma", data = heats) #model converges without error

res0heats <- simulateResiduals(fittedModel = PM_heats_model0)
plot(res0heats) #residuals a bit skewed
testZeroInflation(res0heats) #also lookign good

PM_heats_model0
summary(PM_heats_model0)
Anova(PM_heats_model0)

plot(X.D~PartMort,heats)

#can we cram in interaction?
PM_heats_model1 <- glmmTMB(PartMort ~ X.D*Origin + (1|Raceway), ziformula = ~X.D*Origin , family = "ziGamma", data = heats) #some issues arise, again likely due to partial mortality occurring in all symbiodinium dominated genets in heat...

res1heats <- simulateResiduals(fittedModel = PM_heats_model1)
plot(res1heats) #good
testZeroInflation(res1heats) #also lookign good

PM_heats_model1
summary(PM_heats_model1)
Anova(PM_heats_model1) #cannot estimate


#what about among family levels within origin?
pmOri <- glht(PM_heats_model0, linfct = mcp(Origin_PostHoc = "Tukey"))
summary(pmOri) #E-H-SI sig diff from E-B B - does this make sense?

sub2<-heats %>% #just positive PM
  filter(PartMort>0)

ggplot(sub2, aes(Origin_PostHoc, PartMort)) + geom_boxplot() #yes

#now lets print table of coefficients
PMsum<-summary(PM_dif_model0)
PMsum<-summary(PM_heats_model0)

pxt(lme4::formatVC(PMsum$varcor$cond),"Random effect variances")
pxt(coef(PMsum)$cond,"Conditional fixed effects")
pxt(coef(PMsum)$zi,"Conditional zero-inflation effects")

#upload code to: https://latexbase.com/d/9eb8490b-dd46-48b1-89ea-258aebec6647 to get tables

```

Daily Growth Rate

```{r}

hist(traitsAllCrossSym$DailyGrowthRate_cor, breaks=20)

DGR_dif_model <- lmer(DailyGrowthRate_cor ~ X.D+Origin_PostHoc*Treatment  + (1|Genet) + (1|Raceway), data=traitsAllCrossSym) #default is gaussian(identity)

test <- simulateResiduals(fittedModel = DGR_dif_model)
plot(test) #not terrible, but not perfect...is model too complex?

DGR_dif_model2 <- lmer(DailyGrowthRate_cor ~ X.D+Origin_PostHoc*Treatment  + (1|Raceway), data=traitsAllCrossSym) #

test2 <- simulateResiduals(fittedModel = DGR_dif_model2)
plot(test2) #not really better

summary(DGR_dif_model2) #what if we decide this model is good enough...

anova(DGR_dif_model2,ddf="Kenward-Roger") #only affect of origin


traitsAllCrossSym %>% 
  ggplot(aes(x = Origin_PostHoc, y = DailyGrowthRate_cor)) + 
  geom_boxplot() #verify origin effect

rand(DGR_dif_model2) #no effect of raceway...maybe don't need random effect? do see big outliers in biparental

Ori <- glht(DGR_dif_model2, linfct = mcp(Origin_PostHoc = "Tukey"))
summary(Ori) 


```

## Now quantile normalize and export focal trait lists for gwas

No need to regress for other traits at this point, will rely on relatedness matrix to deal with cross origin. Will have to return to full dataset though as there is no a priori reason to subset to just crosses, can include few other randos since we have genomic data for all

```{r}

str(traitsFiltered_SymTypeSubset)

inv.normalize <- function(x) { #this would also tolerate NAs
  return( qnorm( (rank(x, na.last = "keep") - 0.5) / sum(!is.na(x))))}


ByGenetD<-traitsFiltered_SymTypeSubset %>%                   #get ordered vector of genets
  group_by(Genet) %>%
  summarise(meanD=mean(X.D,na.rm=TRUE))

ByGenetD$normD<-inv.normalize(ByGenetD$meanD)

  

ByGenetMort<-traitsFiltered_SymTypeSubset %>%                    #get ordered vector of genets
  filter(Treatment=="Heat") %>%
  group_by(Genet) %>%
  summarise(propDead = mean(Mortality))

ByGenetMort$normMort<-inv.normalize(ByGenetMort$propDead)

ByGenetSAc<-traitsFiltered_SymTypeSubset %>%                    #get ordered vector of genets
  filter(Treatment=="Control") %>%
  group_by(Genet) %>%
  summarise(meanSAc = mean(SA_dif_cor,na.rm=TRUE))

summary(ByGenetSAc)
hist(ByGenetSAc$meanSAc) #very nice distribution - can quantile norm here

ByGenetSAc$normSAc<-inv.normalize(ByGenetSAc$meanSAc)

plot(ByGenetSAc$meanSAc~ByGenetSAc$normSAc)#good, reflects rankings pretty well

hist(ByGenetSAc$normSAc) #and is nice and normal

ByGenetSAh<-traitsFiltered_SymTypeSubset %>%                    #get ordered vector of genets
  filter(Treatment=="Heat") %>%
  group_by(Genet) %>%
  summarise(meanSAh = mean(SA_dif_cor,na.rm=TRUE))

ByGenetSA<-full_join(ByGenetSAc,ByGenetSAh) 

summary(ByGenetSA)

ByGenetSA %>%
  filter(meanSAh==0)

ByGenetSA$percSA<-(ByGenetSA$meanSAh/ByGenetSA$meanSAc) #create value representing mean proportion of SA growth maintained in heat by a given genet. 1 = no change by treatment

summary(ByGenetSA)

ByGenetSA %>%
  filter(percSA>1) #few genets grew remarkably more in heat

ByGenetSA$percSAnorm<-inv.normalize(ByGenetSA$percSA)

plot(ByGenetSA$percSAnorm~ByGenetSA$percSA)#good, still reflects rankings pretty well

hist(ByGenetSA$percSAnorm) #and is still nice and normal

ByGenetA<-full_join(ByGenetD,ByGenetMort) 

ByGenet<-full_join(ByGenetA,ByGenetSA)

head(ByGenet)
tail(ByGenet)

p1<-ByGenet %>%
  ggplot(aes(x=meanD)) +
    geom_histogram( color="grey60", alpha=0.5, position = 'identity') +
  labs(x="Mean Proportion Durusdinium") #really non-normal, more approximates binomial OR ordinal if we bin intermediate sym types as in stats analysis

p2<-ByGenet %>%
  ggplot(aes(x=normD)) +
    geom_histogram( color="grey60", alpha=0.5, position = 'identity') +
  labs(x="Quantile Normalized Proportion Durusdinium")#transform makes it normal, but appears to randomly assign ones to some positive value and that is not really appropriate either IMO

ByGenet %>%
  ggplot(aes(x=propDead)) +
    geom_histogram( color="grey60", alpha=0.5, position = 'identity') #really more like a categorical ordinal 

ByGenet %>%
  ggplot(aes(x=normMort)) +
    geom_histogram( color="grey60", alpha=0.5, position = 'identity')

p3<-ByGenetD %>%
  ggplot(aes(x = meanD, y = normD)) + 
  geom_point() +
  labs(x="Mean Proportion Durusdinium", y="Quantile Normalized Proportion Durusdinium")

grid.arrange(p1, p2, p3, nrow = 1) #IS transform reasonable

# lets try a plan B which will also be to treat traits as binary and ordinal

ByGenet$D_3bins<-cut(ByGenet$meanD,
                  breaks=c(0,0.1,0.9,1),
                  labels=c('0','1','2')) #defining transitional as more than 10% but less than 90%

ByGenet$D_binary<-cut(ByGenet$meanD, #anything less than 90% is 0
                  breaks=c(0,0.9,1),
                  labels=c('0','1')) 

ByGenet$Mort_bins<-cut(ByGenet$propDead,
                  breaks=c(-0.1,0.25,0.5,0.75,1),
                  labels=c('3','2','1','0'))

ByGenet$Mort_binary<-cut(ByGenet$propDead,
                  breaks=c(-0.1,0.25,1),
                  labels=c('1','0'))

plot(ByGenet$Mort_bins)
plot(ByGenet$Mort_binary)

head(ByGenet)
str(ByGenet)



# fiByGenet# finally, need to read in additional sample ID info to organize genets by seq sample name for GWAS

ids=read.csv("~/Dropbox/CarlsLab/ResearchProjects/NOAA_RuthGates/GenomeAnalysis/RelatednessKey.csv",stringsAsFactors=TRUE) # read in sample name key
nrow(ids)
head(ids)

ids<-ids[,c(1,3,4,5)]

names(ids)<-c("seqID","cross" , "sampleID","Genet")

head(ids)
str(ids)

ids$Genet<-as.factor(ids$Genet)

ByGenetAnnot<-full_join(ids,ByGenet, by="Genet")

head(ByGenetAnnot)
str(ByGenetAnnot)
summary(ByGenetAnnot)

ByGenetAnnot %>%
  filter(is.na(Genet)) #good - just missing wild adult traits

#obtain sample order from vcf/.fam file
order<-c('21073D-03-01_S46_L003',	'21073D-03-02_S75_L005',	'21073D-03-03_S76_L005',	'21073D-03-04_S77_L005',	'21073D-03-05_S47_L003',	'21073D-03-06_S78_L005',	'21073D-03-07_S48_L003',	'21073D-03-08_S79_L005',	'21073D-03-100_S118_L003',	'21073D-03-101_S99_L005',	'21073D-03-102_S100_L005',	'21073D-03-103_S101_L005',	'21073D-03-104_S102_L005',	'21073D-03-105_S119_L003',	'21073D-03-106_S120_L003',	'21073D-03-107_S121_L003',	'21073D-03-108_S122_L003',	'21073D-03-109_S123_L003',	'21073D-03-10_S49_L003',	'21073D-03-110_S124_L003',	'21073D-03-111_S125_L003',	'21073D-03-112_S126_L003',	'21073D-03-113_S127_L003',	'21073D-03-114_S128_L003',	'21073D-03-115_S129_L003',	'21073D-03-116_S130_L003',	'21073D-03-117_S131_L003',	'21073D-03-118_S132_L003',	'21073D-03-119_S133_L003',	'21073D-03-11_S50_L003',	'21073D-03-120_S134_L003',	'21073D-03-121_S135_L003',	'21073D-03-122_S136_L003',	'21073D-03-123_S137_L003',	'21073D-03-124_S138_L003',	'21073D-03-125_S139_L003',	'21073D-03-126_S140_L003',	'21073D-03-127_S141_L003',	'21073D-03-128_S142_L003',	'21073D-03-129_S143_L003',	'21073D-03-12_S80_L005',	'21073D-03-130_S144_L003',	'21073D-03-131_S145_L003',	'21073D-03-132_S146_L003',	'21073D-03-133_S147_L003',	'21073D-03-134_S148_L003',	'21073D-03-135_S149_L003',	'21073D-03-136_S150_L003',	'21073D-03-137_S151_L003',	'21073D-03-138_S152_L003',	'21073D-03-139_S153_L003',	'21073D-03-140_S103_L005',	'21073D-03-141_S104_L005',	'21073D-03-142_S154_L003',	'21073D-03-143_S105_L005',	'21073D-03-144_S155_L003',	'21073D-03-145_S106_L005',	'21073D-03-146_S107_L005',	'21073D-03-147_S156_L003',	'21073D-03-148_S108_L005',	'21073D-03-149_S109_L005',	'21073D-03-14_S51_L003',	'21073D-03-150_S110_L005',	'21073D-03-152_S111_L005',	'21073D-03-153_S129_L006',	'21073D-03-154_S157_L003',	'21073D-03-155_S158_L003',	'21073D-03-156_S159_L003',	'21073D-03-157_S160_L003',	'21073D-03-15_S52_L003',	'21073D-03-16_S53_L003',	'21073D-03-17_S85_L005','21073D-03-18_S54_L003',	'21073D-03-19_S86_L005',	'21073D-03-20_S55_L003',	'21073D-03-21_S81_L005',	'21073D-03-23_S56_L003','21073D-03-24_S57_L003',	'21073D-03-25_S58_L003',	'21073D-03-26_S59_L003',	'21073D-03-27_S60_L003',	'21073D-03-28_S61_L003',	'21073D-03-30_S63_L003',	'21073D-03-31_S64_L003',	'21073D-03-32_S82_L005',	'21073D-03-33_S65_L003',	'21073D-03-34_S66_L003',	'21073D-03-35_S67_L003',	'21073D-03-36_S68_L003',	'21073D-03-37_S69_L003',	'21073D-03-38_S70_L003',	'21073D-03-39_S71_L003',	'21073D-03-40_S72_L003',	'21073D-03-41_S73_L003',	'21073D-03-42_S74_L003',	'21073D-03-43_S75_L003',	'21073D-03-44_S76_L003',	'21073D-03-45_S77_L003',	'21073D-03-46_S78_L003',	'21073D-03-47_S79_L003',	'21073D-03-49_S80_L003',	'21073D-03-51_S81_L003',	'21073D-03-52_S82_L003',	'21073D-03-53_S83_L003',	'21073D-03-54_S84_L003',	'21073D-03-55_S85_L003',	'21073D-03-56_S86_L003',	'21073D-03-57_S87_L003',	'21073D-03-58_S90_L005',	'21073D-03-59_S88_L003',	'21073D-03-60_S89_L003',	'21073D-03-61_S90_L003',	'21073D-03-62_S91_L003',	'21073D-03-63_S92_L003',	'21073D-03-64_S93_L003',	'21073D-03-65_S94_L003',	'21073D-03-66_S91_L005',	'21073D-03-67_S95_L003',	'21073D-03-68_S96_L003',	'21073D-03-69_S97_L003',	'21073D-03-70_S98_L003',	'21073D-03-71_S99_L003',	'21073D-03-72_S100_L003',	'21073D-03-73_S101_L003',	'21073D-03-74_S102_L003',	'21073D-03-75_S103_L003',	'21073D-03-76_S104_L003',	'21073D-03-77_S105_L003',	'21073D-03-78_S106_L003',	'21073D-03-79_S92_L005',	'21073D-03-80_S107_L003',	'21073D-03-81_S108_L003',	'21073D-03-82_S93_L005',	'21073D-03-83_S94_L005',	'21073D-03-84_S109_L003',	'21073D-03-85_S110_L003',	'21073D-03-86_S111_L003',	'21073D-03-87_S95_L005',	'21073D-03-88_S112_L003',	'21073D-03-89_S113_L003',	'21073D-03-90_S96_L005',	'21073D-03-91_S97_L005',	'21073D-03-92_S98_L005',	'21073D-03-93_S114_L003',	'21073D-03-94_S115_L003',	'21073D-03-95_S116_L003',	'21073D-03-96_S117_L003',	'21073D-03-97_S161_L003',	'21073D-03-98_S162_L003',	'21073D-03-99_S163_L003',	'21073D-02-29_S165_L003')

order<-as.data.frame(order)

rownames(order)<-order$order

rownames(ByGenetAnnot)<-ByGenetAnnot$seqID

ByGenetAnnot<-ByGenetAnnot[rownames(order),]

ByGenetAnnot$seqID

names(ByGenetAnnot)

summary(ByGenetAnnot) #down to missing traits for just one genet

write.table(x = ByGenetAnnot, 
              file = "Phenotypes.tab", 
              row.names = F, 
              col.names = F, 
              quote = F, 
              sep = '\t')

write.table(x = ByGenetAnnot, 
              file = "PhenotypesHeader.tab", 
              row.names = F, 
              col.names = TRUE, 
              quote = F, 
              sep = '\t')

ByGenetAnnot %>%
  filter(is.na(D_binary))

traitsFiltered_SymTypeSubset %>%
  filter(Genet=="358") #yes, missing in OG dataset okay

```
